---
title: "homework5"
output: github_document
---
In the ISLR book, read section 6.1.3 “Choosing the Optimal Model” and section 5.1 “Cross-Validation”. Extend and convert the attached effective-df-aic-bic-mcycle.R R script into an R markdown file that accomplishes the following tasks.
```{r}
library('MASS') ## for 'mcycle'
library('manipulate') ## for 'manipulate'
library(tidyverse)
library(caret)
```
# Question 1: Randomly split the mcycle data into training (75%) and validation (25%) subsets.
```{r}
#load data
mcycle <- mcycle
#split the mcycle data into training (75%) subsets
train<-sample_frac(mcycle, 0.75)
sid<-as.numeric(rownames(train))
# split the mcycle data into validation (25%) subsets
test<-mcycle[-sid,]
```

# Question 2: Using the mcycle data, consider predicting the mean acceleration as a function of time. Use the Nadaraya-Watson method with the k-NN kernel function to create a series of prediction models by varying the tuning parameter over a sequence of values. (hint: the script already implements this)


```{r}
# predicting the mean acceleration
y <- train$accel
x <- matrix(train$times, length(train$times), 1)
```
## k-NN kernel function
```{r}
## x  - n x p matrix of training inputs
## x0 - 1 x p input where to make prediction
## k  - number of nearest neighbors
kernel_k_nearest_neighbors <- function(x, x0, k=1) {
  ## compute distance betwen each x and x0
  z <- t(t(x) - x0)
  d <- sqrt(rowSums(z*z))

  ## initialize kernel weights to zero
  w <- rep(0, length(d))
  
  ## set weight to 1 for k nearest neighbors
  w[order(d)[1:k]] <- 1
  
  return(w)
}
```
## Make predictions using the NW method
```{r}
## y  - n x 1 vector of training outputs
## x  - n x p matrix of training inputs
## x0 - m x p matrix where to make predictions
## kern  - kernel function to use
## ... - arguments to pass to kernel function
nadaraya_watson <- function(y, x, x0, kern, ...) {
  k <- t(apply(x0, 1, function(x0_) {
    k_ <- kern(x, x0_, ...)
    k_/sum(k_)
  }))
  yhat <- drop(k %*% y)
  attr(yhat, 'k') <- k
  return(yhat)
}
```


```{r}
## create a grid of inputs 
x_plot <- matrix(seq(min(x),max(x),length.out=100),100,1)

## make predictions using NW method at training inputs
y_hat_plot <- nadaraya_watson(y, x, x_plot,
  kernel_k_nearest_neighbors)

## plot predictions
plot(x, y, xlab="Time (ms)", ylab="Acceleration (g)")
lines(x_plot, y_hat_plot, col="#882255", lwd=2) 
```


```{r}
## set k from 1 to 10
k_seq=seq(1, 10, by = 1)
## make predictions using NW method at training inputs
for (n in k_seq){
  y_hat_plot <- nadaraya_watson(y, x, x_plot, kernel_k_nearest_neighbors, n)
}
```





# Question 4: With the squared-error loss function, compute and plot the training error, AIC, BIC, and validation error (using the validation data) as functions of the tuning parameter.
```{r}
#Use validation data
y <- test$accel
x <- matrix(test$times, length(test$times), 1)
```
## the squared-error loss function
```{r}
## loss function
## y    - train/test y
## yhat - predictions at train/test x
loss_squared_error <- function(y, yhat)
  (y - yhat)^2
```

## test/train error
```{r}
## y    - train/test y
## yhat - predictions at train/test x
## loss - loss function
error <- function(y, yhat, loss=loss_squared_error)
  mean(loss(y, yhat))
```

## AIC
```{r}
## y    - training y
## yhat - predictions at training x
## d    - effective degrees of freedom
aic <- function(y, yhat, d)
  error(y, yhat) + 2/length(y)*d
```

## BIC
```{r}
## y    - training y
## yhat - predictions at training x
## d    - effective degrees of freedom
bic <- function(y, yhat, d)
  error(y, yhat) + log(length(y))/length(y)*d
```

```{r}
## Compute effective df using NW method
## y  - n x 1 vector of training outputs
## x  - n x p matrix of training inputs
## kern  - kernel function to use
## ... - arguments to pass to kernel function
effective_df <- function(y, x, kern, ...) {
  y_hat <- nadaraya_watson(y, x, x,
    kern=kern, ...)
  sum(diag(attr(y_hat, 'k')))
}
```

```{r}
## how does k affect shape of predictor and eff. df using k-nn kernel ?

   ## make predictions using NW method at training inputs
   y_hat <- nadaraya_watson(y, x, x,
     kern=kernel_k_nearest_neighbors, k=1)
   edf <- effective_df(y, x, 
     kern=kernel_k_nearest_neighbors, k=1)
   aic_ <- aic(y, y_hat, edf)
   bic_ <- bic(y, y_hat, edf)
   y_hat_plot <- nadaraya_watson(y, x, x_plot,
     kern=kernel_k_nearest_neighbors, k=1)
   plot(x, y, xlab="Time (ms)", ylab="Acceleration (g)")
   legend('topright', legend = c(
     paste0('eff. df = ', round(edf,1)),
     paste0('aic = ', round(aic_, 1)),
     paste0('bic = ', round(bic_, 1))),
     bty='n')
   lines(x_plot, y_hat_plot, col="#882255", lwd=2) 
```



# Question 4: For each value of the tuning parameter, Perform 5-fold cross-validation using the combined training and validation data. This results in 5 estimates of test error per tuning parameter value.

## 5-fold cross-validation of knnreg model

```{r}
## create five folds
set.seed(100)
mcycle_flds  <- createFolds(mcycle$accel, k=5)
print(mcycle_flds)
sapply(mcycle_flds, length) 
```
## 5 estimates of test error 
```{r}
cvknnreg <- function(kNN = 10, flds=mcycle_flds) {
  cverr <- rep(NA, length(flds))
  for(tst_idx in 1:length(flds)) { ## for each fold
    
    ## get training and testing data
    mcycle_trn <- mcycle[-flds[[tst_idx]],]
    mcycle_tst <- mcycle[ flds[[tst_idx]],]
    
    ## fit kNN model to training data
    knn_fit <- knnreg(accel ~ times,
                      k=kNN, data=mcycle_trn)
    
    ## compute test error on testing data
    pre_tst <- predict(knn_fit, mcycle_tst)
    cverr[tst_idx] <- mean((mcycle_tst$accel - pre_tst)^2)
  }
  return(cverr)
}
```

# Question 5: Plot the CV-estimated test error (average of the five estimates from each fold) as a function of the tuning parameter. Add vertical line segments to the figure (using the segments function in R) that represent one “standard error” of the CV-estimated test error (standard deviation of the five estimates from each fold).

# CV-estimated test error
```{r}
## Compute 5-fold CV for kNN = 1:20
cverrs <- sapply(1:20, cvknnreg)
print(cverrs) ## rows are k-folds (1:5), cols are kNN (1:20)
cverrs_mean <- apply(cverrs, 2, mean)
cverrs_sd   <- apply(cverrs, 2, sd)
## Plot the results of 5-fold CV for kNN = 1:20
plot(x=1:20, y=cverrs_mean, 
     ylim=range(cverrs),
     xlab="'k' in kNN", ylab="CV Estimate of Test Error")
segments(x0=1:20, x1=1:20,
         y0=cverrs_mean-cverrs_sd,
         y1=cverrs_mean+cverrs_sd)
best_idx <- which.min(cverrs_mean)
points(x=best_idx, y=cverrs_mean[best_idx], pch=20)
abline(h=cverrs_mean[best_idx] + cverrs_sd[best_idx], lty=3) 
```



# Question 6:Interpret the resulting figures and select a suitable value for the tuning parameter.

