---
title: "homework7"
output: github_document
---


Goal: Get started using Keras to construct simple neural networks
Work through the "Image Classification" tutorial on the RStudio Keras website.
Use the Keras library to re-implement the simple neural network discussed during lecture for the mixture data (see nnet.R). Use a single 10-node hidden layer; fully connected.
Create a figure to illustrate that the predictions are (or are not) similar using the 'nnet' function versus the Keras model.
(optional extra credit) Convert the neural network described in the "Image Classification" tutorial to a network that is similar to one of the convolutional networks described during lecture on 4/15 (i.e., Net-3, Net-4, or Net-5) and also described in the ESL book section 11.7. See the !ConvNet tutorial on the RStudio Keras website.

`

```{r}
library(keras)
library(tensorflow)
```



```{r}
# load data
fashion_mnist <- dataset_fashion_mnist()
c(train_images, train_labels) %<-% fashion_mnist$train
c(test_images, test_labels) %<-% fashion_mnist$test
```

```{r}
class_names = c('T-shirt/top',
                'Trouser',
                'Pullover',
                'Dress',
                'Coat', 
                'Sandal',
                'Shirt',
                'Sneaker',
                'Bag',
                'Ankle boot')
```


```{r}
#exploration
dim(train_images)
```
```{r}
dim(train_labels)
```
```{r}
train_labels[1:20]
```
```{r}
dim(test_labels)
```
```{r}
library(tidyr)
library(ggplot2)

image_1 <- as.data.frame(train_images[1, , ])
colnames(image_1) <- seq_len(ncol(image_1))
image_1$y <- seq_len(nrow(image_1))
image_1 <- gather(image_1, "x", "value", -y)
image_1$x <- as.integer(image_1$x)

ggplot(image_1, aes(x = x, y = y, fill = value)) +
  geom_tile() +
  scale_fill_gradient(low = "white", high = "black", na.value = NA) +
  scale_y_reverse() +
  theme_minimal() +
  theme(panel.grid = element_blank())   +
  theme(aspect.ratio = 1) +
  xlab("") +
  ylab("")
```
```{r}
train_images <- train_images / 255
test_images <- test_images / 255
```
```{r}
par(mfcol=c(5,5))
par(mar=c(0, 0, 1.5, 0), xaxs='i', yaxs='i')
for (i in 1:25) { 
  img <- train_images[i, , ]
  img <- t(apply(img, 2, rev)) 
  image(1:28, 1:28, img, col = gray((0:255)/255), xaxt = 'n', yaxt = 'n',
        main = paste(class_names[train_labels[i] + 1]))
}
```
```{r}
#build model
model <- keras_model_sequential()
model %>%
  layer_flatten(input_shape = c(28, 28)) %>%
  layer_dense(units = 128, activation = 'relu') %>%
  layer_dense(units = 10, activation = 'softmax')
# compile model
model %>% compile(
  optimizer = 'adam', # how model is updated based on data and loss
  loss = 'sparse_categorical_crossentropy', 
  metrics = c('accuracy') 
)

# train model
model %>% fit(train_images, train_labels, epochs = 5, verbose = 2) # ~89% accuracy
```
```{r}
# evaluate error
score <- model %>% evaluate(test_images, test_labels, verbose = 0)
cat('Test loss:', score['loss'], "\n")
#test accuracy
cat('Test accuracy:', score['accuracy'], "\n") # less accurate than train data, overfit
```
```{r}
#make prediction
predictions <- model %>% predict(test_images)
predictions[1, ] 
```
```{r}
which.max(predictions[1, ])
```

```{r}
class_pred=model %>% predict(test_images) %>%k_argmax()
class_pred[1:20]
```
```{r}
test_labels[1]
```

```{r}
par(mfcol=c(5,5))
par(mar=c(0, 0, 1.5, 0), xaxs='i', yaxs='i')
for (i in 1:25) { 
  img <- test_images[i, , ]
  img <- t(apply(img, 2, rev)) 
  # subtract 1 as labels go from 0 to 9
  predicted_label <- which.max(predictions[i, ]) - 1
  true_label <- test_labels[i]
  if (predicted_label == true_label) {
    color <- '#008800' 
  } else {
    color <- '#bb0000'
  }
  image(1:28, 1:28, img, col = gray((0:255)/255), xaxt = 'n', yaxt = 'n',
        main = paste0(class_names[predicted_label + 1], " (",
                      class_names[true_label + 1], ")"),
        col.main = color)
}
```


```{r}
img <- test_images[1, , , drop = FALSE]
dim(img)
```
```{r}
predictions <- model %>% predict(img)
predictions
```
```{r}
# subtract 1 as labels are 0-based
prediction <- predictions[1, ] - 1
which.max(prediction)
```
```{r}
class_pred <- model %>% predict(img) %>%k_argmax()
class_pred
```











