---
title: "homework4"
output: github_document
---

Question 4

 When the number of features p is large, there tends to be a deteri- oration in the performance of KNN and other local approaches that perform prediction using only observations that are near the test observation for which a prediction must be made. This phenomenon is
known as the curse of dimensionality, and it ties into the fact that curse of di- non-parametric approaches often perform poorly when p is large. We will now investigate this curse.

(a) Suppose that we have a set of observations, each with measure- ments on p = 1 feature, X. We assume that X is uniformly (evenly) distributed on [0, 1]. Associated with each observation is a response value. Suppose that we wish to predict a test obser- vation’s response using only observations that are within 10 % of the range of X closest to that test observation. For instance, in order to predict the response for a test observation with X = 0.6, we will use observations in the range [0.55,0.65]. On average, what fraction of the available observations will we use to make the prediction?

$$ \int_{0.05}^{0.95}10dx+\int_{0.05}^0(100x+5)dx+\int_{0.95}^1(105−100x)dx=9+0.375+0.375=9.75 $$

(b) Now suppose that we have a set of observations, each with measurements on p = 2 features, X1 and X2. We assume that (X1, X2) are uniformly distributed on [0, 1] × [0, 1]. We wish to predict a test observation’s response using only observations that are within 10 % of the range of X1 and within 10 % of the range of X2 closest to that test observation. For instance, in order to predict the response for a test observation with X1 = 0.6 and X2 = 0.35, we will use observations in the range [0.55, 0.65] for X1 and in the range [0.3,0.4] for X2. On average, what fraction of the available observations will we use to make the prediction?

$$ (9.75\%×9.75\%)=0.950625\% $$

(c) Now suppose that we have a set of observations on p = 100 fea- tures. Again the observations are uniformly distributed on each feature, and again each feature ranges in value from 0 to 1. We wish to predict a test observation’s response using observations within the 10 % of each feature’s range that is closest to that test observation. What fraction of the available observations will we use to make the prediction?
$$  (9.75)^p=(9.75)^{100}≃0% $$



(d) Using your answers to parts (a)–(c), argue that a drawback of KNN when p is large is that there are very few training obser- vations “near” any given test observation.

$$\displaystyle \lim_{p \to \infty} (9.75\%)^p=0$$


(e) Now suppose that we wish to make a prediction for a test observation by creating a p-dimensional hypercube centered around the test observation that contains, on average, 10 % of the training observations. For p = 1,2, and 100, what is the length of each side of the hypercube? Comment on your answer. Note: A hypercube is a generalization of a cube to an arbitrary number of dimensions. When p
 = 1, a hypercube is simply a line segment, when p = 2 it is a square, and when p= 100 it is a 100-dimensional cube.

When P =1: 
$$l=0.1$$

When P =2 
$$l=0.1^{1/2}$$

When P =100
$$l=0.1^{1/100}$$





Question 10
This question should be answered using the Weekly data set, which is part of the ISLR2 package. This data is similar in nature to the Smarket data from this chapter’s lab, except that it contains 1, 089 weekly returns for 21 years, from the beginning of 1990 to the end of 2010.
(a) Produce some numerical and graphical summaries of the Weekly data. Do there appear to be any patterns?

```{r}
library(ISLR)
```
```{r}
summary(Weekly)
```
```{r}
require(corrplot)
cor=cor(Weekly[,1:8])
corrplot(cor,diag = FALSE,method="number")
```


(b) Use the full data set to perform a logistic regression with Direction as the response and the five lag variables plus Volume as predictors. Use the summary function to print the results. Do any of the predictors appear to be statistically significant? If so, which ones?
```{r}
Weekly_fit<-glm(Direction~Lag1+Lag2+Lag3+Lag4+Lag5+Volume, data=Weekly,family=binomial)
summary(Weekly_fit)
```
From the coefficients charts, we are able to tinf that log 2 is statistically significant.

(c) Compute the confusion matrix and overall fraction of correct predictions. Explain what the confusion matrix is telling you about the types of mistakes made by logistic regression.
```{r}
Weekly_prid <- predict(Weekly_fit, type = "response")
Weekly_prid_glm <- ifelse(Weekly_prid>0.5,"up","down")
Weekly_table=table(Weekly_prid_glm,Weekly$Direction)
Weekly_table
#percentage of correct predictions
percentage_correct_predictions=(Weekly_table[1,1]+Weekly_table[2,2])/(Weekly_table[1,1]+Weekly_table[1,2]+Weekly_table[2,1]+Weekly_table[2,2])
percentage_correct_predictions
#percentage of Up prediction
percentage_of_Up_prediction=Weekly_table[2,2]/(Weekly_table[2,2]+Weekly_table[1,2])
percentage_of_Up_prediction
#percentage of Down prediction
percentage_of_Down_prediction=Weekly_table[1,1]/(Weekly_table[1,1]+Weekly_table[2,1])
percentage_of_Down_prediction
```
We can find that  the model predicted the overall weekly correct rate 56.11% of the time. Speciafically, the model correctly predicts the Up and Down trends. The percentage of Up correct prediction is 92.07% and the percentage of Down correct prediction is just 11.16%

(d) Now fit the logistic regression model using a training data period from 1990 to 2008, with Lag2 as the only predictor. Compute the confusion matrix and the overall fraction of correct predictions for the held out data (that is, the data from 2009 and 2010).
```{r}
train_data <- (Weekly$Year <2009)
fit_lg2 <- glm(Direction ~ Lag2, data = Weekly, family = binomial, subset = train_data)
pred_lg2 <- predict(fit_lg2, Weekly[!train_data, ], type = "response")
pred_lg2_glm <- ifelse(pred_lg2>0.5,"up","down")
Weekly_table=table(pred_lg2_glm , Weekly$Direction[!train_data])
Weekly_table
#percentage of correct predictions
percentage_correct_predictions=(Weekly_table[1,1]+Weekly_table[2,2])/(Weekly_table[1,1]+Weekly_table[1,2]+Weekly_table[2,1]+Weekly_table[2,2])
percentage_correct_predictions
#percentage of Up prediction
percentage_of_Up_prediction=Weekly_table[2,2]/(Weekly_table[2,2]+Weekly_table[1,2])
percentage_of_Up_prediction
#percentage of Down prediction
percentage_of_Down_prediction=Weekly_table[1,1]/(Weekly_table[1,1]+Weekly_table[2,1])
percentage_of_Down_prediction
```
We can find that  the model predicted the overall weekly correct precnetage increase to 62.5% from 56.11% of the time. Speciafically, the model correctly predicts the Up and Down trends. The percentage of Up correct prediction decrease to 91.80% from 92.07% and the percentage of Down correct prediction increase to 20.93% from 11.16%

(e) Repeat (d) using LDA.
```{r}
library(MASS)
Weeklylda_fit<-lda(Direction~Lag2, data=Weekly,family=binomial, subset=train_data)
Weeklylda_pred<-predict(Weeklylda_fit, Weekly[!train_data, ])
Weekly_table=table(Weeklylda_pred$class, Weekly$Direction[!train_data])
Weekly_table
#percentage of correct predictions
percentage_correct_predictions=(Weekly_table[1,1]+Weekly_table[2,2])/(Weekly_table[1,1]+Weekly_table[1,2]+Weekly_table[2,1]+Weekly_table[2,2])
percentage_correct_predictions
#percentage of Up prediction
percentage_of_Up_prediction=Weekly_table[2,2]/(Weekly_table[2,2]+Weekly_table[1,2])
percentage_of_Up_prediction
#percentage of Down prediction
percentage_of_Down_prediction=Weekly_table[1,1]/(Weekly_table[1,1]+Weekly_table[2,1])
percentage_of_Down_prediction
```
 the result is same with e
(f) Repeat (d) using QDA.
```{r}
Weeklyqda_fit = qda(Direction ~ Lag2, data = Weekly, subset = train_data)
Weeklyqda_pred = predict(Weeklyqda_fit, Weekly[!train_data, ])$class
Weekly_table=table(Weeklyqda_pred , Weekly$Direction[!train_data])
Weekly_table
#percentage of correct predictions
percentage_correct_predictions=(Weekly_table[1,1]+Weekly_table[2,2])/(Weekly_table[1,1]+Weekly_table[1,2]+Weekly_table[2,1]+Weekly_table[2,2])
percentage_correct_predictions
#percentage of Up prediction
percentage_of_Up_prediction=Weekly_table[2,2]/(Weekly_table[2,2]+Weekly_table[1,2])
percentage_of_Up_prediction
#percentage of Down prediction
percentage_of_Down_prediction=Weekly_table[1,1]/(Weekly_table[1,1]+Weekly_table[2,1])
percentage_of_Down_prediction
```
We can find that  the model predicted the overall weekly correct precnetage decrease to 58.66%. Speciafically, the model correctly predicts the Up and Down trends. The percentage of Up correct prediction increase to 100% but the percentage of Down correct prediction decrease to 0
(g) Repeat (d) using KNN with K = 1.
```{r}
library(class)
Weekly_train_knn <- as.matrix(Weekly$Lag2[train_data])
Weekly_test_knn <- as.matrix(Weekly$Lag2[!train_data])
Weekly_train_Direction <- Weekly$Direction[train_data]
set.seed(2)
pred.knn <- knn(Weekly_train_knn, Weekly_test_knn , Weekly_train_Direction, k = 1)
Weekly_table=table(pred.knn,Weekly$Direction[!train_data])
Weekly_table
#percentage of correct predictions
percentage_correct_predictions=(Weekly_table[1,1]+Weekly_table[2,2])/(Weekly_table[1,1]+Weekly_table[1,2]+Weekly_table[2,1]+Weekly_table[2,2])
percentage_correct_predictions
#percentage of Up prediction
percentage_of_Up_prediction=Weekly_table[2,2]/(Weekly_table[2,2]+Weekly_table[1,2])
percentage_of_Up_prediction
#percentage of Down prediction
percentage_of_Down_prediction=Weekly_table[1,1]/(Weekly_table[1,1]+Weekly_table[2,1])
percentage_of_Down_prediction
```
We can find that  the model predicted the overall weekly correct precnetage is still 50%. Speciafically, the model correctly predicts the Up and Down trends. The percentage of Up correct prediction is 50.82% and the percentage of Down correct prediction decrease to 48.84%

(i) Which of these methods appears to provide the best results on this data?

Overall, both original Logistic regression and LDA models have the best results on this data

(j) Experiment with different combinations of predictors, including possible transformations and interactions, for each of the methods. Report the variables, method, and associated confusion matrix that appears to provide the best results on the held out data. Note that you should also experiment with values for K in the KNN classifier.
```{r}
# Logistic regression with Lag2:Lag5
fit_log2_log5 <- glm(Weekly$Direction ~ Lag2:Lag5+Lag2, data = Weekly, family = binomial, subset = train_data)
fit_log2_log5_pred <- predict(fit_log2_log5,Weekly[!train_data, ], type = "response")
fit_log2_log5_pred_glm <- ifelse(fit_log2_log5_pred>0.5,"up","down")
Weekly_table=table(fit_log2_log5_pred_glm, Weekly$Direction[!train_data])
Weekly_table
#percentage of correct predictions
percentage_correct_predictions=(Weekly_table[1,1]+Weekly_table[2,2])/(Weekly_table[1,1]+Weekly_table[1,2]+Weekly_table[2,1]+Weekly_table[2,2])
percentage_correct_predictions
#percentage of Up prediction
percentage_of_Up_prediction=Weekly_table[2,2]/(Weekly_table[2,2]+Weekly_table[1,2])
percentage_of_Up_prediction
#percentage of Down prediction
percentage_of_Down_prediction=Weekly_table[1,1]/(Weekly_table[1,1]+Weekly_table[2,1])
percentage_of_Down_prediction

```
```{r}
#ida with Lag2:Lag5
lda_log2_log5 = lda(Direction~Lag2:Lag5+Lag2, data=Weekly,subset= train_data)
lda_log2_log5_pred= predict(lda_log2_log5, Weekly[!train_data, ])
Weekly_table=table(lda_log2_log5_pred$class, Weekly$Direction[!train_data])
Weekly_table
#percentage of correct predictions
percentage_correct_predictions=(Weekly_table[1,1]+Weekly_table[2,2])/(Weekly_table[1,1]+Weekly_table[1,2]+Weekly_table[2,1]+Weekly_table[2,2])
percentage_correct_predictions
#percentage of Up prediction
percentage_of_Up_prediction=Weekly_table[2,2]/(Weekly_table[2,2]+Weekly_table[1,2])
percentage_of_Up_prediction
#percentage of Down prediction
percentage_of_Down_prediction=Weekly_table[1,1]/(Weekly_table[1,1]+Weekly_table[2,1])
percentage_of_Down_prediction
```
Here we can find the accuracy improved a little bit, but not too much

